Popularity, Mixed Matchings, and Self-duality

### 2017-04-03 CO750-ApproximationAlgorithmsTopics/assignments/assignment_03.tex
3b starting correct argument

Now we change $(WTAP)$ so that the requirement is $(V, E\cup \bar{L})$ is $2$-vertex-connected.

>>>>

Now we change $(WTAP)$ so that the requirement is $(V, E\cup \bar{L})$ is $2$-vertex-connected. Since links between adjacent vertices do not help $2$-vertex-connectivity we may assume there are no such links, otherwise just discard them before starting the algorithm. We will modify the algorithm $\cA$ as follows, the big changes are that we take a leaf to be the root, and we ``shorten" the downward arcs, and we add ``cross" arcs:
\item Select an arbitrary leaf $r \in V$ as the root. Orient $E$ towards $r$, and set the $w(a) = 0$ for all the resulting arcs $a$.
	\item If neither $u$ nor $v$ are each others' ancestor then  add to $G$ the arcs $(u,v)$ and $(v,u)$. Set $w(u,v) = w(v,u) = c(uv)$. We also say the arcs $(u,v)$ and $(v,u)$ ``correspond to the link $uv$. Note that at most four arcs are generated between this step and the previous step for each link.
	\item For each downward arc ($p$ is an ancestor of $u$) $(p,u)$ do: if $p$ is not a leaf let $c_p$ be the child of $p$  that succeeds $p$ on the path to $u$ (otherwise let $c_p = p$). Replace the arc $(p,u)$ with the arc $(c_p, u)$ with the same weight, and same corresponding link as $(p,u)$.
\item Output the set $\bar{L} \subseteq L$ such that $\ell \in \bar{L}$ iff there exists an arc in $R$ that corresponds to $\ell$.
\begin{lemma}\label{lemma:3b-poly}
As argued in lemma \ref{lemma:3a-poly} the algorithm $\cA$ runs in polynomial time, it is clear that the additional processing runs in polynomial time. To verify feasibility suppose for a contradiction that the graph $H := (V, E \cup \bar{L})$ has a vertex $p$ whose deletion disconnects the graph. Let $g$ be the parent of $p$. The vertex $g$ exists since we choose $r$ to be a leaf and hence $p \neq r$. Let $u_1, \dots, u_k$ be the children of $p$. For each $v \in V$, let $C(v,p)$ denote the connected component of $H-p$ containing vertex $v$. Let $A := \{u \in \{u_1, \dots, u_k\} : u \in C(g,p)\}$ be the set of children of $p$ which are connected to $g$ after removing $p$. We claim that $A \neq \emptyset$.
Clearly $r \in C(g,p)$, as are all ancestors of $p$. Now since $R$ is an $out$-arborescence there exists a directed path from $r$ to $\bigcup_{i=1}^k u_i$. The arc such a path uses to enter $\bigcup_{i=1}^k u_i$ is a downward arc, from either $p$ or  some ancestor of $p$, $a$, to a descendant of $p$, $d$. Let $u$ be a child of $p$ such that $d \in C(u,p)$. If the arc is of the form $(a,d)$ then the corresponding link  in $\bar{L}$ connects a vertex of $C(g,p)$ to some descendant of a child $u$. Hence $C(g,p)$ is connected to $C(u)$. That is $u \in A$ and hence $A \neq \emptyset$. In the case the arc is of the form $(p,d)$, then by step $2(c)$, the arc $(p,d)$ arose from a downward arc $(g,d)$. Thus there is a link in $\bar{L}$ connecting a vertex of $C(g,p)$ to $d$. Thus $C(g)$ is connected to $C(u)$. Therefore our claim that $A$ is not empty holds.
 Let $\bar{A} = \{u_1, \dots, u_k\} \backslash A$. By our contradiction assumption, $\bar{A}\neq \emptyset$ as otherwise removing $p$ does not disconnect $H$. Now, again since $R$ is an $out$-arborescence there exists a directed path from $r$ to $\bigcup_{u \in \bar{A}} C(u)$. The directed arc which enters $\bigcup_{u \in \bar{A}} C(u)$ is either a downward arc or a cross arc arising from step $2(b)$. In the case it is a downward arc, by the same argument as the previous paragraph the corresponding link in $\bar{L}$ connects some $C(v, p)$ for $v \in \bar{A}$ to some $C(s,p)$ for $s \in A \cup \{g\}$. But this contradicts the definition of $\bar{A}$. In the case of a cross arc, the head is necessarily in $C(g,p)$ and the tail in some $C(v,p)$ for $v \in \bar{A}$. Thus, as in the previous case, the corresponding link in $\bar{L}$ connects some $C(v, p)$ for $v \in \bar{A}$ to some $C(s,p)$ for $s \in A \cup \{g\}$. In either case a contradiction is reached since the corresponding components to vertices in $\bar{A}$ cannot be connected to $C(g,p)$ by definition. Therefore no vertex exists whose deletion disconnects the graph.
